\section{Introduction}

%First part - Issue of interest for your paper
%
%talk about monetary models and their failure. Cite Meese and Rogoff
%
Many approaches tried to model FX-rates since the start of a free-floating system after Bretton Woods in the 1970s. Monetary models, linking macroeconomics variables to FX-rates were the first to emerge and due to their compelling theoretical intuition soon became the reference in the field. Nonetheless, the enthusiasm for such elegant models was diminished by two seminal works. Meese and Rogoff (1983a, \cite{MeeseRogoffa}) and Meese and Rogoff (1983b \cite{MeeseRogoffb}) demonstrated how none of the different monetary models could beat a random walk without a drift in modeling FX-rates out of sample. After the evidence was presented the opinion divided into four distinct diverging groups. Some argued for a sample issue and confined the problem to the specific historical period, others argued for the limit of the linear models used by the authors to approach monetary models, others argued for a fail of monetary models and their underlying assumptions and, finally, others started to approach the issue in an innovative way turning towards micro-based models that aim to capture the complexity of market information asymmetries and investors heterogeneity \footnote{See Balliu and King (2005, \cite{BalliuKing}) for a general survey.}.

% Second Part - What we do in a nutshell
While all the above-mentioned different reactions contributed to a further understanding of FX-rates properties we question whether the runaway from linear monetary models is justified given the methodological approach presented in the studies of Meese and Rogoff, the small sample of use and the lack of a cointegration analysis later formalized by Granger and Engle (1987,  \cite{EngleGranger}). 
%Third Part - Motivation. Why is it important? Who cares?
% 
% Avoid overengineering and overcomplication. Provide simple setting
% Understand if macroeconomics variables are cointegrated at monthly lag frequency.
%
This methodological study stands therefore as a revision of the out of sample failure of foreign exchange rates testing whether macroeconomics variables are indeed incapable of explaining FX-rates movements before exploring more nuanced fields such as nonlinear and micro-based modeling. The hope is therefore to avoid over-engineering and over-complication of the issue going back to the fundamentals.

%Fourth Part - What we do in more details
Specifically, our study aims to explore the above at a monthly lag frequency on a self-developed data set for the recent period ranging from 1986 to 2006. This period allows for a consistent estimation of all of the necessary parameters given the chosen models. Moreover, it avoids modeling the most recent financial crises where structural breaks are present in the series.\\
As in Meese and Rogoff (1983a and 1983b, \cite{MeeseRogoffa,MeeseRogoffb}) we explore the out of sample forecasting perfomance of foreign exchange rates for three major monetary models. Firstly, the monetary models presented by Frenkel (1976, \cite{Frenkel1976}) and  Mussa\ (1976, \cite{Mussa1976}) claiming a link among monetary mass differentials, interest rates differentials, output gap and FX-rates. Secondly the sticky price model where prices adjust sluggishly to macroeconomics shocks, exchange rates overshoot and inflation rates differential determine the long run behaviour of FX-rates (Dornbush 1976, \cite{Dornbusch}). Finally, a portfolio balance model claiming for the importance of monetary flows and net current account differentials as a key determinant for FX-rates changes as in Hooper and Morton (1982, \cite{HooperMorton1982}). \\
Based on such monetary models five different stochastic models are analyzed in order to forecast FX-rates. Three of them are presented in the seminal work of Meese and Rogoff (1983a, \cite{MeeseRogoffa}) and represent the benchmark to verify whether the results of the paper hold for the time frame of interest. These models are a simple random walk forecast, an OLS estimation without inclusion of lagged terms and a vector autoregressive model. On the top of these models we analyze the performance of transfer models discussed in Montgomery and Weatherby (1980, \cite{MontgomeryWeatherby}) and the vector error correction model representation of the cointegration relationship between macroeconomic variables and FX-rates proposed by Granger and Engle (1987,  \cite{EngleGranger}).

%Fifth Part - Main Findings
If our results are in line with the seminal papers of Meese and Rogoff in the case of univariate models, the picture resulting from a multivariate fit
contradicts the results of the reference papers. We do not find enough evidence to support the claim of an underperformance of multivariate
linear models in comparison to a simple random walk model. Moreover, we find a strong evidence for cointegration among macroeconomic series and
FX-rates for all of the analyzed country series. This suggests a long term stable relation between the fundamentals and the FX-rates rejecting
the claim of Meese and Rogoff about a weak relation between the two. Finally, we do find that a vector error correction model representation of the cointegration
relation between macroeconomic series and FX-rates outperforms the out-of-sample forecast of FX-rates at all lags and for all of the country series. This proof further
strengthen the hypothesis that a link between monetary models and FX-rates does indeed exist at monthly lag frequency given a long enough sample period that allows
for a reliable multivariate model estimation.

%Seventh part - This paper is organized as follows
The paper continues as follows. Section 2 outlines the methodological approach and describes the techniques used for
the empirical analysis and the out-of-sample forecast. Section 3 continues by introducing the dataset used and the chosen proxies to capture the
macroeconomic fundamentals of interest, Section 4 reports the main results of the empirical analysis and Section 5 concludes the findings.