\bibliographystyle{unsrt}  
%\bibliography{references}  %%% Remove comment to use the external .bib file (using bibtex).
%%% and comment out the ``thebibliography'' section.


%%% Comment out this section when you \bibliography{references} is enabled.
\begin{thebibliography}{1}

\bibitem{BalliuKing}
 Bailliu, J.  and M. King 2005
\newblock What Drives Movements in Exchange Rates?
\newblock In {\em  Bank of Canada Review, \url{https://www.bankofcanada.ca/wp-content/uploads/2010/06/bailliu.pdf}.}

\bibitem{Dornbusch}
Dornbusch, R. Morton 1976
\newblock Expectations and Exchange Rate Dynamics
\newblock In {\em Journal of Political Economy, 84 (6):}, pages 1161-76.

\bibitem{MontgomeryWeatherby}
Montgomery D. and G. Weatherby 1980
\newblock Modeling and Forecasting Time Series Using Transfer Function and Intervention Methods.
\newblock In {\em A I I E Transactions, Taylor \& Francis 12 (4):}, pages 289-307.

\bibitem{EngleGranger}
Engle, R.F. and C.W.J. Granger 1983
\newblock Co-integration and error correction: Representation, estimation and testing
\newblock In {\em Econometrica 55 (2):}, pages 251¡V276.

\bibitem{Frenkel1976}
Frenkel, J. 1976
\newblock A Monetary Approach to the Exchange Rate: Doctrinal Aspects and Empirical Evidence.
\newblock In {\em Scandinavian Journal of Economics, 78 (2):}, pages 200¡V24.

\bibitem{HooperMorton1982}
Hooper, P. and J. Morton 1982
\newblock Fluctuations in the dollar: A model of nominal and real exchange rate determination.
\newblock In {\em Journal of International Money and Finance, 1 (1):}, pages 39-56.

\bibitem{MeeseRogoffa}
Meese, R.A. and K. Rogoff 1983
\newblock Empirical exchange rate models of the seventies: Do they fit out of sample?
\newblock In {\em Journal of International Economics, 14 (1-2):}, pages 3-24.

\bibitem{MeeseRogoffb}
Meese, R.A. and K. Rogoff 1983
\newblock The Out-of-Sample Failure of Empirical Exchange Rate Models: Sampling Error or Misspecification?
\newblock In {\em Exchange Rates and International Macroeconomics, University of Chicago Press},  ISBN: 0-226-26250-2 14, pages 67 - 112.

\bibitem{Mussa1976}
Mussa, M. 1976
\newblock The Exchange Rate, the Balance of Payments, and Monetary and Fiscal Policy under a Regime of Controlled Floating.
\newblock In {\em Scandinavian Journal of Economics, 78 (2):}, pages 229¡V48.

\end{thebibliography}