---
title: "Linear Modelling"
output: rmarkdown::github_document
---

Useful Libraries
```{r}
## String manipulation library using regualr expressions

library(stringr) ## to make use of regular expressions to parse strings

## To operate with times series objects
library(xts)

## To estimate transfer function 
library(TSA)

## VAR packages

library(vars)  ## to estimate VAR lag with exogenous variables
library(tsDyn) ## to estimate the VAR model and perform the out of sample fit based on rolling window.

## Import self defined functions
source("0 - functions.R")
```

Specify a function to clean the environment and do it
```{r}
clean <- function() 
{
  ENV <- globalenv()
  ll <- ls(envir = ENV)
  ll <- ll[!(ll %in% c("data", "data_univariate", "clean", "import", "write", "evalue", "acf_plots",
                       "adfTest", "seasonality", "suspectFrequency", "spa"))]
  rm(list = ll, envir = ENV)
}

clean()
```

===========
Data Import
===========

```{r}
## Insert your path here
path <- "/Users/Marco Hassan/Exchange-Rates-Modelling/Data_&_Code/Written\ Data/clean_stationary.csv"

data <- read.csv(path)

data <- as.xts(data[,-1], order.by = as.Date(data[,1]))
```


================================
Univariate times series analysis
================================

This section starts with the analysis of the cleaned dataset.

In a first step I will try to check whether the Meese-Rogoff paradox holds in the analyzed time period. I will work with the USA as the benchmark series and apply the different macroeconomics models to linearly estimate the FX_rates.

General model:

EX = const. + b1 M_USFR + b2 UN_USFR + b3 TB_USFR + b4 CPI_USFR + b5(TB_US) + b6 TB_FR

I will test the different models:

(i) Frenkel Bilson -> b4, b5, b6 = 0

(ii) Dornbusch-Frankel -> b5, b6 = 0

(iii) Hooper Morton -> no constraint


As in the case of Meese and Rogoff forecasting will be performed at one, three, six and twelwe months.

Cite Meese and Rogoff in the thesis: 
The purpose of considering multiple forecast horizons in this type of experiment is to see whether the structural models do better than time series models in the long run, when adjustment due to lags and/or a serially correlated error term has taken place. Of course, when lags and serial correlation are fully incorporated into the structural models, a consistently estimated true structural model will outpredict a time series model at all
horizons in a large sample.

===============
Random walk fit
===============

Random walk without drift prediction using rolling series
```{r}
## In the other models I will use 3/4 of the sample to train the data and the rest to predict. Be consistent to facilitate comaprison at a later stage.
three_fourth <- round(nrow(data)*(3/4))

## to create new series transform in data frame as before

data2 <- as.data.frame(data)

for(iPred in c(1,3,6,12)) 
  { 
    helper <- c("JP", "UK", "CH")
    for(jCountry in helper)
    {
    data2[, paste0("rwPred" , iPred, "_", jCountry, "US")] <- NA
    }
  } 

data2 <- as.xts(data2, order.by = index(data))## transform back to xts
data <- data2
rm(data2)
```

Fit the Random Walk prediction
```{r}
for(iPred in c(1,3,6,12)) 
  { 
    helpCountry <- c("JP", "UK", "CH")
    for(jCountry in helpCountry)
    {
        for(kRow in (three_fourth+1):nrow(data)) 
        {
         if ((iPred != 1) & (kRow+(iPred-2) == nrow(data))) break
         data[kRow+(iPred-1), paste0("rwPred" , iPred, "_", jCountry, "US")] <- 
           data[kRow-1, paste0("EX_", jCountry, "US")] 
        }
    }
  } 
```


==============================
Macroeconomic structural model
==============================

In contrast to Meese and Rogoff I decided to make use of the ARIMAX model here. This will allow the introduction of lagged exogenous variables and its implementation is favoured compared to the long lagged AR model used by Meese and Rogoff to my mind.

Specify rolling prediction
```{r}
## trasform to data-frame to add columns
data2 <- as.data.frame(data)

## create prediction columns

for(iPred in c(1,3,6,12)) 
  { 
    helper <- c("JP", "UK", "CH")
    for(jCountry in helper)
    {
    data2[, paste0("strPred" , iPred, "_", jCountry, "US")] <- NA
    }
  } 
```

Get exogenous variables for each of the series of interest
```{r}
## get exogenous parameters
for(country in c("CH", "UK", "JP"))
{
  assign(paste0("interest", country), 
         !str_detect(colnames(data2), paste0("^EX|^M1|^M1+\\D", country)) &  ## Using M3
          str_detect(colnames(data2), paste0("\\D", country)) & 
         !str_detect(colnames(data2), "^rw|^str"))
}
```

### Transfer function fit

```{r}
## transform to ts object ot leverage on the window function
data2 <- as.ts(data2, frequency =12)

## in sample observations
in_sample <- three_fourth;
  
## Get in sample time frame
st <- tsp(data2)[1] + (in_sample-2); 
## tsp(data2)[1] gets the start of the time frame; the second block adds the in_sample time window
```


Test restricted models
```{r}
for( country in c("JP", "CH", "UK") )
  {
    ## Specify the arimax that fits at best
    nd_pars<-expand.grid(ar=0:2,diff=0:1,ma=0:2)
    nd_aic<-rep(0,nrow(nd_pars))
    help <- get(paste0("interest",country))
    xshort <- window(data2, start = tsp(data2)[1]+ (1-1), end=st + (1-1))
    
    for(j in seq(along=nd_aic))
    {
      nd_aic[j]<-AIC(arimax(xshort[,paste0("EX_", country, "US")],
                            unlist(nd_pars[j,1:3]), 
                            xreg = xshort[, help],
                            method = 'ML'))
    }
    
    ## save the best fit
    order <- nd_pars[which.min(nd_aic),]
    
    ## General structural the model
    fit1 <- arimax(xshort[, paste0("EX_", country, "US")], order = unlist(order), 
                         xreg = xshort[, help], method = 'ML')
    
    ## Frenkson Bilson test
    xshort2 <- xshort[, help]
    fit2 <- arimax(xshort[, paste0("EX_", country, "US")], order = unlist(order), 
                         xreg = xshort2[, -(4:5)], method = 'ML')
    
    ## Dornbusch Model test
    fit3 <- arimax(xshort[, paste0("EX_", country, "US")], order = unlist(order), 
                         xreg = xshort2[, -5], method = 'ML')
    
    ## Frenkson Bilson
    cat("\nCountry "); cat(country); cat(":\n")
    cat("Frenkson Bilson w\\o Current account and Inflation\n")
    print(waldtest(fit1, fit2))
    
    ## Dornbusch Model
    cat("Dornbusch Model w\\o Current account\n")
    print(waldtest(fit2, fit3))
  }
```

Working well just last observation cannot be computed due to the lack of exogenous observations.
```{r}
file_con <- file("station_arimax.txt")
sink(file_con, "out")
for( LAG in c(1, 3, 6, 12) ) 
{
  cat("this is lag: "); cat(LAG); cat("\n")
  for( country in c("CH", "JP", "UK") )
  {  
    cat("Country: "); cat(country); cat("\n")
    for( i in 1:(nrow(data2)-in_sample) )
    {
      xshort <- window(data2, start = tsp(data2)[1]+ (i-1), end=st + (i-1))
      xnew <- window(data2, start = st + i, end = (st +i)+(LAG-1))
    
      ## Specify the arimax that fits at best
      nd_pars<-expand.grid(ar=0:2,diff=0:1,ma=0:2)
      nd_aic<-rep(0,nrow(nd_pars))
      help <- get(paste0("interest",country))
      
      for(j in seq(along=nd_aic))
      {
        nd_aic[j]<-AIC(arimax(xshort[,paste0("EX_", country, "US")],
                              unlist(nd_pars[j,1:3]), 
                              xreg = xshort[, help],
                              method = 'ML'))
      }
    
      ## save the best fit
      order <- nd_pars[which.min(nd_aic),]
      cat(paste0(order))
      cat("\n")
      
    
      fit1 <- arimax(xshort[, paste0("EX_", country, "US")], order = unlist(order), 
                     xreg = xshort[, help], method = 'ML')
      
      ## Create an exogenous vector and specify the structur in the case of lag == 1
      ## to keep the column structure in a consistent way to fit the model
      if ( LAG == 1) {new_exogenous <- t(xnew[, help])} else new_exogenous <- xnew[, help]
      
      fcast1 <- predict(fit1, n.ahead = LAG, newxreg = new_exogenous)
      ## breeak prima qui.
      ## so that you have always 12 out of sample
      
      if(index(tail(xshort,1)) == nrow(data2)-LAG) break 
      data2[in_sample+(LAG-1)+i, 
            paste0("strPred", LAG, "_", country, "US")] <- fcast1$pred[1]
    }
  }
}
sink()
close(file_con)
```


## OLS fit
```{r}
## trasform to data-frame to add columns
data2 <- as.data.frame(data2)

## create prediction columns

for(iPred in c(1,3,6,12)) 
  { 
    helper <- c("JP", "UK", "CH")
    for(jCountry in helper)
    {
    data2[, paste0("OLSPred" , iPred, "_", jCountry, "US")] <- NA
    }
} 

## transform to ts object ot leverage on the window function
data2 <- as.ts(data2, frequency =12)

## in sample observations
in_sample <- three_fourth;
  
## Get in sample time frame
st <- tsp(data2)[1] + (in_sample-2); 
## tsp(data2)[1] gets the start of the time frame; the second block adds the in_sample time window

## get exogenous parameters
for(country in c("CH", "UK", "JP"))
{
  assign(paste0("interest", country), 
         !str_detect(colnames(data2), paste0("^EX|^M1|^M1+\\D", country)) &  ## Using M3
          str_detect(colnames(data2), paste0("\\D", country)) & 
         !str_detect(colnames(data2), "^rw|^str|^var|^OLS"))
}
```

Test restricted models
```{r}
for( country in c("JP", "CH", "UK") )
  {
    xshort <- window(data2, start = tsp(data2)[1]+ (1-1), end=st + (1-1))
    help <- get(paste0("interest",country))
    
    ## General structural the model
    fit1 <- arimax(xshort[, paste0("EX_", country, "US")], 
                   xreg = xshort[, help], method = 'ML')
    
    ## Frenkson Bilson test
    xshort2 <- xshort[, help]
    fit2 <- arimax(xshort[, paste0("EX_", country, "US")], 
                   xreg = xshort2[, -(4:5)], method = 'ML')
    
    ## Dornbusch Model test
    fit3 <- arimax(xshort[, paste0("EX_", country, "US")],
                   xreg = xshort2[, -5], method = 'ML')
    
    ## Frenkson Bilson
    cat("Country "); cat(country); cat(":\n")
    cat("Frenkson Bilson w\\o Current account and Inflation\n")
    print(waldtest(fit1, fit2))
    
    ## Dornbusch Model
    cat("Dornbusch Model w\\o Current account\n")
    print(waldtest(fit2, fit3))
  }
```


Working well.
```{r}
for( LAG in c(1, 3, 6, 12) ) 
{
  for( country in c("JP", "CH", "UK") )
  {  
    for( i in 1:(nrow(data2)-in_sample) )
    {
      xshort <- window(data2, start = tsp(data2)[1]+ (i-1), end=st + (i-1))
      xnew <- window(data2, start = st + i, end = (st +i)+(LAG-1))
      
      help <- get(paste0("interest",country))
    
      fit1 <- arimax(xshort[, paste0("EX_", country, "US")], 
                     xreg = xshort[, help], method = 'ML')
      
      ## Create an exogenous vector and specify the structur in the case of lag == 1
      ## to keep the column structure in a consistent way to fit the model
      if ( LAG == 1) new_exogenous <- t(xnew[, help])
      else new_exogenous <- xnew[, help]
      
      fcast1 <- predict(fit1, n.ahead = LAG, newxreg = new_exogenous)
      if(index(tail(xshort,1)) == nrow(data2)-LAG) break 
      ## so that you have always 12 out of sample
      data2[in_sample+(LAG-1)+i, 
            paste0("OLSPred", LAG, "_", country, "US")] <- fcast1$pred[1]
    }
  }
}
```

Convert back to xts
```{r}
data2 <- as.data.frame(data2)
data2 <- as.xts(data2, order.by = index(data))## transform back to xts
data <- data2
```

Clean working space
```{r}
clean()
data2 <- data
```

Save results into a csv
```{r}
write(data, "linear_pred.csv")
```


============================
Autocorrelation of the Error
============================

Check at autocorrelation of error series to see whethere the residuals of the models are white noise of whether there is systematic information that was not captured.
```{r}
# number rolling prediction per forecasting time frame
for(i in c(1,3,6,12))
  {
  assign(paste0("numprediction", i), sum(!is.na(data[, paste0("rwPred", i, "_JPUS")])))
  }

# Plot acf and test for white noise
for ( iLag in c(1, 3, 6, 12) )
{
  for( country in c("CH", "JP", "UK") )
  {
    par(mfrow = c(3,1))
    for( method in c("rw", "str", "OLS") )
    {
      error <- tail(data[, paste0("EX_", country, "US")],
                          evalue(paste0("numprediction", iLag))
                          ) - 
                na.trim(data[, paste0(method, 
                                    "Pred", iLag,"_", 
                                    country, "US")
                            ]
                        )
      acf(error, main = paste("Pred", method, iLag, country))
      
      ## Back up with quantitative Ljung Box measure
      for( jLag in c(5,20,50) )
      {
      print(paste("Pred", method, iLag, country, "Ljung", jLag))
      print(Box.test(error, lag = jLag, type = "Ljung-Box"))
      }
      
    }
  }
}
```

Write the data
```{r}
write(data, "linear_pred.csv")
```


