---
title: "Level Linear Modelling"
output: rmarkdown::github_document
---

Libraries used
```{r}

## String manipulation library using regualr expressions

library(stringr) ## to make use of regular expressions to parse strings

## xts to manipulate times series objects
library(xts)

## For applying Johansen test for cointegration
library(urca)

## For performing ARIMAX evaluation.
library(TSA)

## VAR packages

library(vars)  ## to estimate VAR lag with exogenous variables
library(tsDyn) ## to estimate the VAR model and perform the out of sample fit based on rolling window.
library(TSA)

## Import functions from the relating project
source("0 - functions.R")
```


Specify a function to clean the environment and do it
```{r}
clean <- function() 
{
  ENV <- globalenv()
  ll <- ls(envir = ENV)
  ll <- ll[!(ll %in% c("data", "data_univariate", "clean", "import", "write", "evalue", "acf_plots",
                       "adfTest", "seasonality", "suspectFrequency", "spa"))]
  rm(list = ll, envir = ENV)
}

clean()
```

Import datasets
```{r}
## Insert your path here
path <- "/Users/Marco Hassan/Exchange-Rates-Modelling/Data_&_Code/Written\ Data/rawdata.csv"
data <- read.csv(path)
data <- as.xts(data[,-1], order.by = as.Date(data[,1]))

## Insert your path here
path <- "/Users/Marco Hassan/Exchange-Rates-Modelling/Data_&_Code/Written\ Data/linear_pred.csv"
data_univariate <- read.csv(path)
data_univariate <- as.xts(data_univariate[,-1], order.by = as.Date(data_univariate[,1]))

```

===============================
Vector Autoregressive Modelling
===============================

```{r}
data2 <- as.data.frame(data)
endogenous <- c("EX_UKUS", "EX_JPUS", "EX_CHUS")
exogenous <- !str_detect(colnames(data2), "^M1|^rw|^EX|^str|^OLS")
data_exogenous <-  data2[, exogenous]
```

Create columns for inserting prediction results
```{r}
for(iPred in c(1,3,6,12)) 
  { 
    helper <- c("JP", "UK", "CH")
    for(jCountry in helper)
    {
    data2[, paste0("varPred" , iPred, "_", jCountry, "US")] <- NA
    }
  } 
```

Iportant reminder of the predict_rolling function computations
This function allows to check the out-of sample forecasting accuracy by
estimating the model on a sub-sample of the original, then making
\code{nroll} forcasts of horizont \code{n.ahead}, each time by updating the
sample. In other words, with a given model estimated on 100 observations, the
function will estimate it on say 90 first obs (\code{nroll=10}), generate a
say 1 step-ahead \code{n.ahead=1} from obs 90, then using true value 91,
92,... till full sample.

Rolling prediction using VAR
```{r}
# number rolling prediction per forecasting time frame
for(i in c(1,3,6,12))
  {
  assign(paste0("numprediction", i), sum(!is.na(data_univariate[, paste0("rwPred", i, "_JPUS")])))
  }

## Estimate the rolling window forecast
for(j in c("UK", "JP", "CH")) 
  {
    for(i in c(1,3,6,12))
    {
    lags <- VARselect(data2[, endogenous], lag.max = 7,
          type = "none", exogen = data_exogenous) 
    
    min_lag <- min(lags$selection)

    ## one lag the best
    model <- lineVar(cbind(data2[, endogenous],data_exogenous) , 1, 
                     lag = min_lag,
                     include = "none", 
                     model = "VAR", I = "diff",
                     estim = "ML", LRinclude = "none")
            
    parsed_text1 <- evalue(paste0("numprediction", i))
    assign(paste0("endRow",i), nrow(data2) +1 - parsed_text1)
    assign(paste0("pred_model", i), 
           predict_rolling(model, nroll = parsed_text1, n.ahead = i, 
                           refit.every = 1))
    parsed_text2 <- evalue(paste0("pred_model", i))
    parsed_text3 <- evalue(paste0("endRow", i))
    if(j == "UK") list_member = 1
    if(j == "JP") list_member = 2
    if(j == "CH") list_member = 3
    data2[parsed_text3:nrow(data2), 
         paste0("varPred", i, "_", j, "US")] <- parsed_text2$pred[list_member]
    }
}
```

Convert back to xts
```{r}
data2 <- as.xts(data2, order.by = index(data))## transform back to xts
data <- data2
clean()
```

============================================
Estimation via Vector Error Correction Model 
============================================

Notice here you are running a different model compared to the VAR above. Above you tested a model where the endougenous variables were the exogenous exchange rates were assumed to be possibly endougenous and the maacroeconomic variables were assumed to be exogenous. 

Now the model is DIFFERENT as you are actually testing whether the structural macroeconomic variables are endogenouusly correlated witht eh exchange rates of interest and testing for forecasting superior predictive ability given the stable relation between two of the variables.

Create columns for inserting prediction results
```{r}
## trasform to data-frame to add columns
data2 <- as.data.frame(data)
for(iPred in c(1,3,6,12)) 
  { 
    helper <- c("JP", "UK", "CH")
    for(jCountry in helper)
    {
    data2[, paste0("vecmPred" , iPred, "_", jCountry, "US")] <- NA
    }
  } 
```

Create a data frame containing the cointegrated series for the japan series
```{r}
## Specify in Sample dataset
three_fourth <- round(nrow(data)*(3/4))
```

Rolling prediction using VECM - make use of the same rolling_prediction function
```{r}
# number rolling prediction per forecasting time frame
for(i in c(1,3,6,12))
{
  assign(paste0("numprediction", i), sum(!is.na(data_univariate[, paste0("rwPred", i, "_JPUS")])))
}
## Estimate the rolling window forecast
for( jCountry in c("CH", "JP", "UK") )
{
  if(jCountry == "UK")
  {
     sample <- cbind(
           data2[,paste0("EX_", jCountry, "US")], 
           data2[,paste0("M3_", "US", jCountry)], 
           data2[,paste0("UN_", "US", jCountry)], 
           data2[,paste0("CPI_", "US", jCountry)],
           data2[,paste0("T_", "US", jCountry)]) 
    
  }
  
  else 
  {
     sample <- cbind(
           data2[,paste0("EX_", jCountry, "US")], 
           data2[,paste0("M3_", "US", jCountry)], 
           data2[,paste0("TB_", "US", jCountry)],
           data2[,paste0("UN_", "US", jCountry)], 
           data2[,paste0("CPI_", "US", jCountry)],
           data2[,paste0("T_", "US", jCountry)]) 
  }
  
  ## Create a data frame containing the cointegrated series for the japan series
  assign(paste0("data_", jCountry),  sample)
  
  ## Select Lag length and specify model
  LAG <- min(VARselect(evalue(paste0("data_", jCountry)), 
                       lag.max = 7, type = "none")$selection
             ) 
  
  model <- tsDyn::VECM(evalue(paste0("data_", jCountry)),
                       lag = LAG, estim = "ML", LRinclude = 'none'
                      )
  
  for(i in c(1,3,6,12))
  {
    parsed_text1 <- evalue(paste0("numprediction", i))
    assign(paste0("endRow",i), nrow(data2) +1 - parsed_text1)
    assign(paste0("pred_model", i), 
           predict_rolling(model, nroll = parsed_text1, n.ahead = i, refit.every = 1))
    parsed_text2 <- evalue(paste0("pred_model", i))
    parsed_text3 <- evalue(paste0("endRow", i))
    data2[parsed_text3:nrow(data2), 
          paste0("vecmPred", i, "_", jCountry, "US")] <- parsed_text2$pred[1]
  }
}
```

Convert back to xts
```{r}
data2 <- as.xts(data2, order.by = index(data))## transform back to xts
data <- data2
clean()

data2 <- as.data.frame(data)
data2 <- sapply(data2, diff)

interest <- str_detect(colnames(data2), "^var|^vecm")
data <- data.frame(data_univariate, data2[, interest])
data <- as.xts(data, order.by = index(data_univariate))
```


Write the data
```{r}
write(data, "linear_pred.csv")
```
