---
title: "R Notebook"
output: rmarkdown::github_document
---

=========
Libraries
=========

```{r}
## Import self defined functions
source("0 - functions.R") ## for this section especially important is the SPA function that returns the bootstrapped interval for the superior preddiction ability hypothesis.

## To handle xts times series objects.
library(xts)

## To check at partition splits
library(readr)
library(stringr)

## To get latex tables
library(xtable)

## For performing model confidence set
library(MCS)
```

===========
Import Data
===========
```{r}
path <- "/Users/Marco Hassan/Exchange-Rates-Modelling/Data_&_Code/Written\ Data/time_stationary_GTS_pred.csv"

data <- read.csv(path)

data <- as.xts(data[,-1], order.by = as.Date(data[,1]))
```


Plot the different out of sample forecasting fits
```{r}
for(jCountry in c("CH", "JP", "UK"))
{
  print(plot(tail(data[,paste0("EX_", jCountry, "US")], 50), ylab="Rates", 
     main="Linear Estimators - 1 Lag Out of Sample",
     grid.col = NA, col = 1))
  print(lines(data[,paste0("rwPred1_", jCountry, "US")],col= "red", lwd = 2))
  print(lines(data[,paste0("OLSPred1_", jCountry, "US")], col = "orange", lwd = 2))
  print(lines(data[,paste0("GTSPred1_", jCountry, "US")], col = "salmon", lwd = 2))
  print(addLegend("topleft", on=1, 
            legend.names = c("Original", "RW", "OLS", "GTS"), 
            lty=c(1, 1), lwd=c(2, 1),
            col=c(1, 2, "orange", "salmon"))) 
}

```


** Step 2 - Quantitative Measure for Prediction Power **

Root Mean Squared Error (MSE)
```{r}
rmseFun <- 
          function(error)
          {
            #error <- Actual - Forecast
            #print(error)
            return( sqrt(sum(error**2)/length(error)) )
          }
```

Mean Absolute Error (MAE)
```{r}
maeFun <- 
            function(error)
            {
              #return( Actual - Forecast )
              return( sum(abs(error))/length(error) )
            }
```

Directional Accuracy (DAC)
```{r}
dacFun <- 
          function(Actual, Forecast, LAG)
          {
            return( mean ( sign( diff(Actual, lag=LAG) ) == 
                           sign( diff(Forecast, lag=LAG) ),
                           na.rm = T
                          ) 
                  )
          }
```


Create Series of Forcasting Errors; repeat for all of the lags.
```{r}

# number rolling prediction per forecasting time frame
for(iLag in c(1,3,6,12))
{
assign(paste0("numprediction", iLag), sum(
                                          !is.na(
                                                  data[, 
                                                      paste0("rwPred", iLag, "_JPUS")
                                                      ]
                                                 )
                                          )
       )
  
}

RMSE <- matrix(NA,4,3, 
               dimnames = list( c("rw", "var", "vecm", "GTS"),
               c("CH", "JP", "UK") )
               )

MAE <- matrix(NA,4,3, 
               dimnames = list( c("rw", "var", "vecm", "GTS"),
               c("CH", "JP", "UK") )
               )

DAC <- matrix(NA,4,3, 
               dimnames = list( c("rw", "vecm", "vecm", "GTS"),
               c("CH", "JP", "UK") )
               )
```


Compute root mean square error and mean absolute error for the different lags and linear estimators
```{r}
errorFun <- 
            function(LAG)
            {
              iRow = 0
              jCol = 1
              for( country in c("CH", "JP", "UK") )
              {
                for( method in c("rw", "str", "OLS", "GTS") )
                {
                  iRow = iRow + 1
                  
                  actual_data <- tail(data[, paste0("EX_", country, "US")],
                                      evalue(paste0("numprediction", LAG))
                                      )
                  
                  forecast_data <- na.trim(data[, paste0(method, 
                                                "Pred", LAG,"_", 
                                                country, "US")
                                               ]
                                           )
                  
                  error_data <- actual_data - forecast_data
                  
                  RMSE[iRow, jCol] <- rmseFun(error_data)
                  
                  MAE[iRow, jCol] <- maeFun(error_data)
                  
                  DAC[iRow, jCol] <- dacFun(actual_data, forecast_data, LAG)
                  
                  if(iRow %% 4 == 0)
                  {
                    jCol = jCol + 1
                    iRow = 0
                  }
                
                }
              }
              
              print(paste("RMSE", LAG))
              RMSE <- xtable(RMSE, digits = c(0,4,4,4))
              #print(RMSE)
              
              print(paste("MAE", LAG))
              MAE <- xtable(MAE, digits = c(0,4,4,4))
              print(MAE)
              
              print(paste("DAC", LAG))
              DAC <- xtable(DAC, digits = c(0,4,4,4))
              #print(DAC)
            }
```

Evalue the function for different lags
```{r}
for( iLAG in c(1,3,6, 12) )
errorFun(iLAG)
```

====================
Model Confidence set
====================

Model Confidence set
```{r}
#Create u a matrix with all different univariate predictions: each column, one model prediction
u <- na.trim(cbind(data$rwPred1_JPUS, data$varPred1_JPUS, data$vecmPred1_JPUS, data$GTSPred1_JPUS))

n<-dim(u)[1]
perf<-matrix(NA,n,4)

## RECALL TO CHANGE ALSO HERE THE COUNTRY WHEN ADAPTING AND DERIVING P-VALUES.
for (j in 1:4)
{
  perf[,j]<- abs(data$EX_JPUS-u[,j]) #MAE 
  #perf[,j] <- abs(data$EX_JPUS-u[,j])**2 #MSE
}

protto <- MCSprocedure(perf, alpha = 0.05, B = 1000, statistic = "TR", verbose = T)

#colMeans(perf) ## Perf is a matrix containing for each column the loss function of one forecasting method.
  
#for (k in 1:4)
#{
#  print( paste("Benchmark model", colnames(u[,k]))	)
#  d<-rep(0,5000)
#  s<-1
#  for (i in 1:500) ## iterate 10 times with spa 100 bootstrapped values each. Total = 1000.
#  { 
#     e<-spa(per=perf,column_benchmark_model=k,
#            num_predmodels=4,number_predictions=n,q=0.25,iter=100,periodogram=T)
#     d[s:(s+99)]<-e$stat.boos
#     s<-s+100
#  }
#  print(mean((d>e$stat))) ## fraction 
#}
```

=============================
Analysis of Partitioned Space
=============================

```{r}
partition_points <- colnames(partitioned_space)
for(country in c("UK"))
{
  message(paste("\n\nFor country", country, "the optimal splits are at:\n"))
  for( LAG in c("1", "3", "6", "12") )
  {
        message(paste("\nFor lag", LAG, "the optimal splits are at:\n"))
    
        prova <- read_file(paste0("Split/OLS/GTS_", country, "_", LAG, ".txt"))
        
        patt <- "(?=Highest).*(?<=[:digit:])"
        
        optimal_part <- str_extract_all(prova, pattern = patt)

        for (partition in c("1", "2", "3", "4", "5"))
        {
            cat(paste("\n", partition, "Optimal Partition:\n"))
            print(sum(str_detect(optimal_part[[1]], partition)))
            
            cat(paste("\nPartition", partition, "splits:\n\n"))
            ## Extract first partition characterisitics 
            patt <- paste("(?<=For partition", partition, "the optimal).*(?<=])")
            
            first_part <- str_extract_all(prova, pattern = patt)
            
            for (method in c("UN", "TB", "M", "CPI", "T_", "time"))
            {
              ## Extract variable
              patt <- paste0("(?=", method, ").*(?=%)")
              
              splitVar <- str_extract_all(first_part[[1]], pattern = patt)
              
              splitVar <- unlist(splitVar)
              
              splitVar2 <- str_detect(first_part[[1]], pattern = patt)
              
              if(sum(splitVar2) == 0)next
              message(paste("Method", method, "splits total:\n"))
              print(sum(splitVar2))
              
              brek_pnts <- paste0("\\D", as.character(seq(0,100, (1/16)*100)))
              
              for(brek in brek_pnts)
              {
                if(sum(str_detect(splitVar, brek)) == 0)next
                cat(paste("Break", brek, "number:\n"))
                print(sum(str_detect(splitVar, brek)))
              }
            }
        }
  }
}
```



