---
title: "Data Cleaning"
output: html_document
---

Useful Libraries
```{r}
## Data cleaning libraries

library(forecast) ## for seasonality adjusting the times series
library(tseries) ## -> check when used
library(TSA) ## for applying FFT to get seasonality frequencies
library(strucchange) ## to perform chow test
library(xts) ## to deal with times series objects.

## Import self defined functions
source("0 - functions.R")
```

Specify a function to clean the environment and do it
```{r}
clean <- function() 
{
  ENV <- globalenv()
  ll <- ls(envir = ENV)
  ll <- ll[!(ll %in% c("data", "clean", "import", "write", "evalue", "acf_plots",
                       "adfTest", "seasonality", "suspectFrequency", "spa"))]
  rm(list = ll, envir = ENV)
}
```


===========
Data import 
===========

```{r}
## Insert your path here
path <- "/Users/Marco Hassan/Exchange-Rates-Modelling/Data_&_Code/Written\ Data/rawdata.csv"

data <- read.csv(path)

data <- as.xts(data[,-1], order.by = as.Date(data[,1]))
```


======================
Descriptive Statistics
======================

This section continues with general descriptive statistics trying to understand the behaviour of the exogenous series and trying to test whether the latter are indeed exogenous.

Before diving into the descriptive analysis of the series, I will try to transform the series into stationary series so that it will possible to analyze the various series throught linear and non-linear times series models.

In order to transform the series into stationary series I will:

    1. Remove the seasonal component.

    2. Detrend the series.
    
    3. Test the series for unit roots.
    
    4. Perform chow-tests to check for structural breaks in the resulting series.


##################
Part 1: Detrending
##################

In a first step I will test the necessity for detrending the series.

Plot the series as a help:
```{r}
plotSeries <- 
              function(i)
              {
              plot(i, main = colnames(i))
              }

lis <- list()
lis <- lapply(data, plotSeries)
for(i in 1:ncol(data))
{
  if(i == 1) 
    par(mfrow = c(3,1))
  else if(((i-1)%%3 == 0) & !(i %in% c(18, 21)))
    par(mfrow = c(3,1))
  plot(lis[[i]])
}
```


Plot proxy for variance as a help:
```{r}
varPlot <- function(i)plot((i- mean(i, na.rm = T))**2, main = colnames(i)) ## notice assumes time invariant mean, which is clearly violated in the series.

lis <- list()
lis <- lapply(data, varPlot)

for(i in 1:ncol(data))
{
  if(i == 1)
    par(mfrow = c(3,1))
  else if(((i-1)%%3 == 0) & !(i %in% c(18, 21)))
    par(mfrow = c(3,1))
  plot(lis[[i]])                  
}
```

The majority of the series are clearly not mean stationary. 

For the exchange rate and the money supply it is difficult to judge based on simple plots, these might could appear to be mean stationary but not volatility stationary. 
```{r}
acf_plots <- function()
{
  for(i in 1:ncol(data)){
    if(i == 1)
      par(mfrow = c(3,1), mar = c(4,3,4,2))
    else if(((i-1)%%3 == 0) & !(i %in% c(18, 21)))
      par(mfrow = c(3,1), mar = c(4,3,4,2))
    acf(na.trim(data[,i]), main = colnames(data[,i]))                 
  }
}

## Plot acf
acf_plots() ## slow decaying autocorrelation points to non-stationarity mean series.
## The slow decay suggests that the series would be mean varyiant based on time shifts.
```

Step 2: Choose detrending solution

(i) Remove yearly trend and yarly seasonality by block differenciating with the yearly series average

Then analyze the following options:

(ii) Detrend via differenciation

(iii) Detrend via linear time trend

(iv) Use a moving average filter, this is a 'low pass' filter since it takes data and removes from it the rapidly fluctuating component  to leave the slowly varying estimated
trend.


Plot new Variables
```{r}
lis <- list()
lis <- lapply(data, plotSeries)
for(i in 1:ncol(data)){
    if(i == 1)
      par(mfrow = c(3,1), mar = c(4,3,4,2))
    else if(((i-1)%%3 == 0) & !(i %in% c(18, 21)))
      par(mfrow = c(3,1), mar = c(4,3,4,2))
  plot(lis[[i]])
}
```


Perform detrending via differentiation
```{r}
data2 <- data ## to play then with original dataset
data3 <- lapply(data, diff)
for(i in 1:ncol(data)){data[,i] <- data3[[i]]}
```

Plot new Variables
```{r}
lis <- list()
lis <- lapply(data, plotSeries)
for(i in 1:ncol(data)){
    if(i == 1)
      par(mfrow = c(3,1), mar = c(4,3,4,2))
    else if(((i-1)%%3 == 0) & !(i %in% c(18, 21)))
      par(mfrow = c(3,1), mar = c(4,3,4,2))
  plot(lis[[i]])
}
```

Plot proxy for variance for the new variables as a help:
```{r}
lis <- list()
lis <- lapply(data, varPlot)
for(i in 1:ncol(data))
{
  if(i == 1)
    par(mfrow = c(3,1), mar = c(4,3,4,2))
  else if(((i-1)%%3 == 0) & !(i %in% c(18, 21)))
    par(mfrow = c(3,1), mar = c(4,3,4,2))
  plot(lis[[i]])                  
}
```


Plot autocorrelation functions
```{r}
## Plot acf
acf_plots()
```


Result: Other possibilitiy is detrending via moving average smoother.

Compute moving average smoother for Trade.
```{r}
#country <- c("JP", "CH", "UK")

#par(mfrow = c(3,1), mar = c(4,4,4,2))
#for(i in country){
#          trend <- ma(data2[, paste0("T_US", i)], 
#                      order = 4, centre = T) ##moving average of quartal
#          data[, paste0("T_US", i)] <- as.ts(data2[, paste0("T_US", i)]) - trend
#          acf(na.trim(data[,paste0("T_US", i)]), main = paste0("T_US", i)) 
#}
```


Check for unit roots in the series performing augmented Dickey-Fuller tests
```{r}
suppressWarnings(lapply(data, adfTest)) ## supress warning to ignore message that true p-value smaller than the one printed
```


Will all of the above corrections the reslut is good and there is strong evidence for stationarity in the series.

Clean workspace and just keep the full dataset
```{r}
data <- na.trim(data)

clean()
```


Step 3: analyze seasonality and correct for it.

Three possible solutions for adjusting for seasonality after identification of the cyclical frequency:

(i) Seasonal differentiation. Similar idea to the general differentiation of a series.

(ii) Subtract the block average of the series in the cycle. Similar idea to moving average smoothers.

(iii) Spectral Analysis and Fourier Modeling of Seasonal Patterns

Let's test now for the presence of seasonal patterns and let's try to identify the cyclical frequency. 

Checking again at autocorrelation plots
```{r}
acf_plots()
```

Perform FFT to check for possible other suspicious seasonality patterns for each series
```{r}
seasonal_pattern <- lapply(data, suspectFrequency)

seasonal_pattern
```

Perform test checking for seasonal unit root at the specified frequency
```{r}
for(iseries in c(1: length(seasonal_pattern))){
  print(colnames(data[,iseries]))
  print(seasonality(data[,iseries], seasonal_pattern[[iseries]][1]))
}
```

specify the double frequency of the series according to the spectral analysis above
```{r}
data <- na.trim(data)
spectral_seasonality <- function(x)
          {
            series <- msts(data[,x], seasonal.periods = c(seasonal_pattern[[x]][1:2]))
            
            taylor.lm <- tslm(series ~ fourier(series, K = c(1, 1)))
            
            data[, x] <<- taylor.lm$residuals
}

for(iseries in colnames(data))
{
  print(iseries)
  spectral_seasonality(iseries)
}
```

Check at new adjusted
```{r}
acf_plots()
```

Perform FFT to check for possible other suspicious seasonality patterns for each series
```{r}
seasonal_pattern <- lapply(data, suspectFrequency)

seasonal_pattern
```

Perform test checking for seasonal unit root at the specified frequency
```{r}
for(iseries in c(1: length(seasonal_pattern))){
  print(colnames(data[,iseries]))
  print(seasonality(data[,iseries], seasonal_pattern[[iseries]][1]))
}
```

Good. No other seasonal unit roots are displayed.

Clean workspace
```{r}
clean()
```


#######################
## Structural breaks ##
#######################

Before starting with the implementation of the linear modelling of the series I perform a chow test to understand the possibility of structural breaks in the sample that would suggest prudence when working with the series as a whole.

This is especially important as the period analyzed involves the outbreak of the dot-com bubble, which called for aggressive monetary policies and might significantly affect the results.
```{r}
breakpnts <- function(x){
  
   ## I test against the alternative that each series can be fitted with a single constant     against the alternative of structural breaks that would call for multiple constants in      the sample.
  
  ## Using CUSUM
  ocus.x <- efp(as.ts(x) ~ 1, type = "OLS-CUSUM")
  plot(ocus.x, main = colnames(x))
}

lapply(data, breakpnts)

```

Write data
```{r}
write(data, "clean_stationary.csv")
```

