
\section*{Appendix A \textendash \ Data}
\label{app:A}
\addcontentsline{toc}{section}{Appendix A \textendash \ Data}

\textit{Series: Unemployment rate.}

Description: Unemployment Rate: Aged 15-64: All Persons.

Source: Main Economic Indicators - complete database, Main Economic Indicators (database),  \url{http://dx.doi.org/10.1787/data-00052-en}, April 13, 2019.

\textit{Series: 3-month Treasury Bills USA, percent annum.}

Description: 3-Month Treasury Bill - Secondary Market Rate 

Source Board of Governors of the Federal Reserve System (US), 3-Month Treasury Bill: Secondary Market Rate [TB3MS], retrieved from FRED, Federal Reserve Bank of St. Louis; \url{https://fred.stlouisfed.org/series/TB3MS}, April 13, 2019.


\textit{Series: Treasury Bills UK.}

Description: Treasury Bills discount rate, percent per annum.

Source: Bank of England, Treasury Bill Discount Rate in the United Kingdom [TBDRUKM], retrieved from FRED, Federal Reserve Bank of St. Louis; \url{https://fred.stlouisfed.org/series/TBDRUKM}, April 13, 2019.


\textit{Series: Treasury Bills JP.}

Description: Interest Rates, Government Securities, Treasury Bills for Japan.

Source: International Monetary Fund, Interest Rates, Government Securities, Treasury Bills for Japan [INTGSTJPM193N], retrieved from FRED, Federal Reserve Bank of St. Louis; \url{https://fred.stlouisfed.org/series/INTGSTJPM193N}, April 13, 2019.

\textit{Series: 3-month LIBOR.}

Description: 3-Month or 90-day Rates and Yields: Interbank Rates

Source: Organization for Economic Co-operation and Development, Main Economic Indicators (database), \url{http://dx.doi.org/10.1787/data-00052-en}, April 13, 2019.

\textit{Series: M3 USA \& JP.}

Description: M3 comprises notes and coins in circulation outside banking corporations; demand and savings deposits, fixed and installment savings deposits, time deposits, and certificates of deposit of households, nonfinancial corporations, local governments, securities companies, Tanshi companies, and some other financial corporations such as securities finance companies with depository corporations in national and foreign currency; and nonresident deposits with banking corporations in national currency. 

Source: International Monetary Fund, M3 for Japan, and US retrieved from FRED, Federal Reserve Bank of St. Louis; \url{https://fred.stlouisfed.org/series/MYAGM3JPM189N} \& \url{https://fred.stlouisfed.org/series/MYAGM3USM052N}, April 13, 2019.

\textit{Series: M3 CHF \& UK.}

Description: M3 comprises notes and coins in circulation outside banking corporations; demand and savings deposits, fixed and installment savings deposits, time deposits, and certificates of deposit of households, nonfinancial corporations, local governments, securities companies, Tanshi companies, and some other financial corporations such as securities finance companies with depository corporations in national and foreign currency; and nonresident deposits with banking corporations in national currency. 

Source: Organization for Economic Co-operation and Development, Main Economic Indicators (database), \url{http://dx.doi.org/10.1787/data-00052-en}, April 13, 2019.

\textit{Series: M1 UK \& JP \& CH.}

Description: M1 consists of: (1) currency outside the U.S. Treasury, Federal Reserve Banks, and the vaults of depository institutions; (2) traveler's checks of nonbank issuers; (3) demand deposits; and (4) other checkable deposits (OCDs), which consist primarily of negotiable order of withdrawal (NOW) accounts at depository institutions and credit union share draft accounts.

Source: Organization for Economic Co-operation and Development, Main Economic Indicators (database), \url{http://dx.doi.org/10.1787/data-00052-en}, April 13, 2019.

\textit{Series: CPI.}

Description: Consumer Price Index: All Items

Source: Organization for Economic Co-operation and Development, Main Economic Indicators (database), \url{http://dx.doi.org/10.1787/data-00052-en}, April 13, 2019.


\textit{Series: Trade.}

Description: Net Trade. Value of Goods.

Source: Organization for Economic Co-operation and Development, Main Economic Indicators (database), \url{http://dx.doi.org/10.1787/data-00052-en}, April 13, 2019.