---
title: "Cointegration in FX-rates"
output:  rmarkdown::github_document
---

Libraries used
```{r}

## String manipulation library using regualr expressions

library(stringr) ## to make use of regular expressions to parse strings

## To perform the adf.test
library(tseries)

## For applying Johansen test for cointegration
library(urca)

## VAR packages

library(vars)  ## to estimate VAR lag with exogenous variables
library(tsDyn) ## to estimate the VAR model and perform the out of sample fit based on rolling window.

## Import functions from the relating project
source("~/Exchange-Rates-Modelling/Master thesis - codice/Code/0 - functions.R")
```


===================================
Import dataset from related project
===================================

Raw series, no deternding nor seasonal adjustment
```{r}
data <- read.csv("/Users/Marco Hassan/Exchange-Rates-Modelling/Master thesis - codice/Written\ Data/rawdata.csv")

data <- as.xts(data[,-1], order.by = as.Date(data[,1]))
```


======================
Descriptive Statistics
======================

This section continues with general descriptive statistics trying to understand the behaviour of the exogenous series and trying to test whether the latter are indeed exogenous.

Get exogenous parameters
```{r}
## get country specific parameters
for(country in c("CH", "UK", "JP"))
{
  assign(paste0("interest", country), 
         !str_detect(colnames(data), paste0("^M1+\\D", country)) & ## with M3
         str_detect(colnames(data), paste0("\\D", country))
        )
}

## Notice that if you perform the above just with the CPI, EX and M1 you find 2 cointegrated vectors at very high significance lag. This means that EX is indeed part of the cointegration relation. 
```

In sample - Three fourth of the dataset
```{r}
three_fourth <- round((3/4)* nrow(data), 0)
```


Johansen cointegration test
```{r}
## transform in data frame to iterate over
data2 <- as.data.frame(data)

for(interest in c("interestJP", "interestCH", "interestUK"))
{
  coint <- ca.jo(data2[1:three_fourth,eval(parse(text=interest))], ecdet = "none", type = "trace")
  print(summary(coint))
}

```


To be sure set up the OLS regression and test for stationarity in the error model
```{r}
fit1 <- lm(EX_JPUS ~ M3_USJP + TB_USJP + UN_USJP + CPI_USJP, data = data2[1:three_fourth,])

error<-ts(residuals(fit1))
adf.test(error) 
```


**Good evidence for the cointegration in the series.**



